---
title: "Social Network of Pen Brands and Customers"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This R Markdown file creates a bipartite social network using brand and customer relationships based on a dataset from the pen industry. The goal is to visualize the connections between brands and their customers.

## Load Required Libraries
```{r libraries}
# Load necessary libraries
# You can install them with install.packages("tidyverse") and install.packages("igraph")
# install.packages(c("tidyverse", "igraph"))  # Uncomment if not already installed
library(tidyverse) # For data manipulation and cleaning
library(igraph) # For building and visualizing network graphs
```

## Load and Inspect the Data
```{r load-data}
# Load your data. Make sure the file path is correct.
# Data is assumed to be semicolon-separated
df <- read_csv("MOT133A_Final_database_V2.csv", col_types = cols(.default = "c"), locale = locale(encoding = "UTF-8"))

# Take a quick look at the structure of the data
glimpse(df)
```

## Data Cleaning
```{r data-cleaning}
# Clean the data: remove rows with missing or empty Brand or Customer, and trim whitespace
df <- df %>%
  filter(!is.na(Brand), Brand != "", !is.na(Customer), Customer != "") %>%
  mutate(
    Brand = str_trim(Brand),
    Customer = str_trim(Customer)
  )

  # Remove non valid UTF-8 characters from all character columns
  df <- df %>%
    mutate(across(where(is.character), ~iconv(., from = "UTF-8", to = "UTF-8", sub = "")))

# Add five new columns with value 0
df <- df %>%
  mutate(
    Budget = 0,
    Entry_level = 0,
    Business = 0,
    Luxury = 0,
    Collector = 0
  )

# Convert Customer column to string (character)
df <- df %>%
  mutate(Customer = as.character(Customer))


# Update the new columns based on Customer values
# Patterns for customer types
budget_pattern <- c("1", "Mass Market", "Budget")
entry_level_pattern <- c("2", "Entry Level", "Entry-level", "Student")
business_pattern <- c("3", "Business", "Professional")
luxury_pattern <- c("4", "Luxury", "Premium")
collector_pattern <- c("5", "Collector", "Luxury Collector")

# Update the new columns based on Customer values
df <- df %>%
  mutate(
    Budget = if_else(grepl(paste(budget_pattern, collapse = "|"), Customer, ignore.case=TRUE), 1, Budget),
    Entry_level = if_else(grepl(paste(entry_level_pattern, collapse = "|"), Customer, ignore.case=TRUE), 1, Entry_level),
    Business = if_else(grepl(paste(business_pattern, collapse = "|"), Customer, ignore.case=TRUE), 1, Business),
    Luxury = if_else(grepl(paste(luxury_pattern, collapse = "|"), Customer, ignore.case=TRUE), 1, Luxury),
    Collector = if_else(grepl(paste(collector_pattern, collapse = "|"), Customer, ignore.case=TRUE), 1, Collector)
  )

  # Merge duplicate rows where Brand and Company are the same
  # For binary columns, take the maximum (1 if present in any duplicate)
  df <- df %>%
    group_by(Brand, Company) %>%
    summarise(
      across(everything(), ~ if(is.numeric(.)) max(., na.rm = TRUE) else first(.)),
      .groups = "drop"
    )

# Delete unnecessary columns
df <- df %>% select(-No, -B2B, -B2C, -Pricing, -Customer)

# View the first few rows of the cleaned data
head(df)
```

## Create an Edge List
```{r create-edge-list}
# Create edges between Brand and each Customer Type

# Gather the customer type columns into long format to create edges
df_edges <- df %>%
  pivot_longer(
    cols = c(Budget, Entry_level, Business, Luxury, Collector),
    names_to = "Customer_Type",
    values_to = "Has_Type"
  ) %>%
  filter(Has_Type == 1) %>%
  select(Brand, Customer_Type)

# Calculate node sizes: Brands by their Market Size
brand_sizes <- df %>%
  select(Brand, `Market Size`) %>%
  distinct() %>%
  mutate(size = as.numeric(`Market Size`)) %>%
  mutate(size = ifelse(is.na(size) | !is.finite(size), 5, size)) %>%
  select(Brand, size)

# For Customer_Type, set to 5 as a constant size
customer_types <- unique(df_edges$Customer_Type)
customer_type_sizes <- tibble(name = customer_types, size = 5)

# Set Color for BIC and Sailor Pen Company
company_colors <- df %>%
  select(Brand, Company) %>%
  distinct() %>%
  mutate(color = case_when(
    Company == "BIC Corporate" ~ "#fdb813",
    Company == "Sailor Pen Company" ~ "#0054a4",
    TRUE ~ "grey"
  ))

# Assign colors to companies
filtered_colors <- company_colors %>%
  select(Brand, color)

# For customer types, set a default color
customer_type_colors <- tibble(name = customer_types, color = "skyblue")

# Combine into a single data frame for node attributes
node_sizes <- bind_rows(
  brand_sizes %>% rename(name = Brand),
  customer_type_sizes
)

node_colors <- bind_rows(
  filtered_colors %>% rename(name = Brand),
  customer_type_colors
)

# View the first few rows of the edge list
head(df_edges)
```

## Build the Network Graph
```{r build-graph}
# Create an undirected graph from the edge list
g <- graph_from_data_frame(df_edges, directed = FALSE)

# Assign types to nodes (Brand = FALSE, Customer = TRUE)
V(g)$type <- bipartite_mapping(g)$type

# Assign sizes to nodes based on the node_sizes data frame
V(g)$size <- ifelse(V(g)$type, 5, node_sizes$size[match(V(g)$name, node_sizes$name)])

# Assign colors to nodes based on the brand_colors mapping
V(g)$color <- ifelse(V(g)$type, "skyblue", node_colors$color[match(V(g)$name, node_colors$name)]) # Default color for brands 
```

## Plot the Graph
```{r plot-graph}
# Plot the bipartite graph with basic styling
plot(
  g,
  vertex.label.cex = 0.3, # Size of labels
  vertex.size = V(g)$size, # Smaller for customers
  vertex.color = V(g)$color, # Color by type
  layout = layout_with_fr # Force-directed layout
)
```

## Community Detection
```{r community-detection, message=FALSE}
# Community Detection (using fast greedy algorithm)
# Only works on connected, undirected graphs
if (is.connected(g)) {
  g_simple <- simplify(g, remove.multiple = TRUE, remove.loops = TRUE)
  comm <- cluster_fast_greedy(g_simple)
  plot(
    comm, g_simple,
    vertex.label.cex = 0.3,
    vertex.size = V(g_simple)$size,
    vertex.color = membership(comm),
    layout = layout_with_fr,
    main = "Community Detection"
  )
} else {
  cat("Graph is not connected; community detection may not be meaningful.\n")
}
```

## Convert to Image
```{r convert-to-image, echo=FALSE, message=FALSE}
# Save the plot as an image
png("pen_brands_network.png", width = 1000, height = 1000, res = 300, pointsize = 6)
plot(
  g,
  vertex.label.cex = .5, # Size of labels
  vertex.size = V(g)$size*2, # Smaller for customers
  vertex.color = V(g)$color, # Color by type
  layout = layout_with_fr # Force-directed layout
)
# Save the community detection plot as an image
dev.off()
png("pen_brands_community.png", width = 1000, height = 1000, res = 300, pointsize = 6)
if (is.connected(g)) {
  g_simple <- simplify(g, remove.multiple = TRUE, remove.loops = TRUE)
  comm <- cluster_fast_greedy(g_simple)
  plot(
    comm, g_simple,
    vertex.label.cex = 0.5,
    vertex.size = V(g_simple)$size*2,
    vertex.color = membership(comm),
    layout = layout_with_fr,
    main = "Community Detection"
  )
} else {
  plot.new()
  text(0.5, 0.5, "Graph is not connected; community detection not shown.")
}
dev.off()
```

## Conclusion
This simple network highlights how pen brands connect with different customers. You can expand this further by including companies, countries, or product types as nodes or attributes.

## Release all resources
```{r cleanup, include=FALSE}
# Clear the environment
rm(list = ls())
```